### 2017-04-11 paper/samplebody-journals.tex
added .gitignore, final pdf update

\quad Yelp is a website which allows users to create an account and review businesses. Our goal is to analyze data from Yelp to gain further insight for business owners to develop a satisfactory experience for their customers. Yelp has a public dataset available on their website with information about users, businesses, and reviews.

\quad Our first objective is to determine if certain attributes are indicative of business success with respect to their average review rating (see sec. 2.3 for revenue relation). If we can find a way to predict whether or not a business will do well later on in its operation, that information will be useful to businesses looking to improve their model.

\quad Our second objective is to analyze the text of reviews for a sentiment. The sentiment analysis will show the emotions conveyed by the reviewer as being pleased or angry at their experience with the business. Using the results of this analysis we hope to be able to determine if certain times of day, city locations, or other factors contribute to the experience of a customer. This information could also contribute to a business model. As an example, if certain locations are more prone to angry customers it could indicate that the people in that area have high standards, the businesses are poorly run, or just that those customers tend to leave angry reviews.

\quad Our last objective is to find recurring problems in businesses. We will analyze review text to find the specific issues that people are having with businesses. If a business has many reviews which all indicate a similar issue, then that is something a business owner would want to be aware of so that they can eliminate the problem.

@@

A previous winner, Hood et al., published a paper on assessing a business?s current rating/review state and future expectations with respect to review numbers on Yelp. The bulk of their work centered around the creation of additional features for the data (e.g. number of similar businesses within 1km, features of subsets of reviews such as average rating and number of unique users). From there they utilized various feature reduction and clustering techniques to build a prediction model. They were able to identify features that offered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as a create model for prediction that outperformed a basic linear regression.

@@

The dataset has many optional attributes. There is an attribute called 'attributes' that is a list of optional attributes. The dataset is inconsistent and cannot be relied upon for each business. For instance, one of these attributes is 'Bike Parking'. This attribute is either true, false, or nonexistent. Speciifcally the fact that this attribute might not be available for multiple businesses makes it unreliable.

\quad In addition, the data is not only restaurants but any kind of business. As a result they have another piece of data which is the categories list. For each business this attribute attempts to classify the business. However some categories are extremely broad and the same type of business can be described in different ways. One example is that there is a bakery, but it?s also categorized as food. Many businesses will sell food, but the category of food will not be in their list of categories and instead it will be labelled with more specific categories like Italian food.

\quad The majority of the data cleaning process will be involved in handling missing data. It is possible to  ll in these missing values with default values like false but this can be dangerous to make assumptions about what businesses have and don?t have.

\quad The Yelp dataset contains up to one million different attributes about businesses from 11 different cities in 4 countries. The majority of these attributes will be unnecessary so there will be a large effort in data reduction. For instance there are photographs included, but our questions does not require image analysis. There also attributes such as postal codes that are unnecessary because latitude and longitude is more accurate.

@@

The Yelp dataset comes from the Yelp Dataset Challenge. This is an ongoing challenge and this is the ninth round. The challenge offers \$5,000 to the winner and the  final submission is an academic paper.

@@

\quad The tools and methods used for this project will include a database for storing and integrating the data, libraries and modules for cleaning and preprocessing, and various tools for transformation, mining, pattern evaluation, and knowledge presentation. Given the json-formatted data, we've opted for using a NoSQL database like MongoDB or Cassandra. Fortunately, there are a number of options, considering our educational status, that will allow large storage volume as well as low latency read/write access. Hosting in such a manner will require a number of Python libraries like Dora and FTFY ('fixes text for you') intended for data cleaning and preprocessing will be used for these purposes. Yelp has already performed a substantial amount of cleaning on the data, so it is likely that minimal additional cleaning will be necessary, save for replacing missing values as mentioned above.

\quad

\quad Numpy and Scipy will be used to transform and mine the data, and other Python modules like Seaborn and Plotly (a Python API for D3) will be used for knowledge visualization and presentation. Finding angry reviews and recurring problems will make heavy use of NLP, so the team will rely on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, and possibly some custom supervised text classifiers trained with the Facebook fastText pre-trained word vectors to perform sentiment analysis on the review text.

@@

\quad Evaluating the sentiment analysis of the review text will likely be simple given that we will be using tools like Weka and Microsoft Sentiment Analysis (see following section) to perform sentiment analysis rather than creating entirely custom algorithms. These tools incorporate helpful metrics to analyze the error of a particular output. Alternatively, we can manually verify the sentiment of a (relatively small) number of reviews by reading them and comparing our human judgements of the sentiment to the output of the tools. For evaluating recurring problems, we will employ a similar approach, except we will group reviews by business and analyze them as a time series to and recurring issues.

\quad For all three of the objective questions, employing association rule frameworks that use support-confidence frameworks will be helpful to and strong associations, while null-invariant measures like the Kulczynski measure or the cosine measure will be helpful to determine how interesting these associations are.

@@

\quad There have been several major changes to the project tools and methods. Initially we planned on using a relational database, but ran into a lot of trouble. The inconsistencies between the user and business data made it too difficult to successfully import it into the database, so instead we decided to import the data straight into Weka. However, the format of the JSON file did not allow a successful import into Weka.

\quad Next we tried to convert the JSON file into CSV format. Once again Weka had issues due to the mismatch in attributes for each business (some had more attributes while others had less).

\quad Instead we opted to use a different tool, RapidMiner. Like Weka, RapidMiner incorporates a extensive machine learning functionality like classification, regression, clustering, and association and frequent itemset mining. RapidMiner is commercial software, but an educational version is available that removes limitations on the free version like only being able to process 10,000 rows at a time. Additionally, RapidMiner is more compatible with the missing data in our dataset and automatically corrects the formatting issues we had when attempting to use Weka. It does not support JSON so we reused the conversion code (found on the team GitHub as convert.py) to convert the data files to CSV.

@@

\quad After importing the CSV formatted version of the data we tested some basic visualization of the data. For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it?s located. Basic graphs like these help refine our question. For example, why does Illinois have significantly lower average stars than others?

@@

\begin{figure}[h]
\includegraphics[width=0.75\textwidth, center]{b_rating_by_state}
\end{figure}

@@

\begin{figure}[!h]
\includegraphics[width=0.75\textwidth, center]{3d_businesses_lat_long}
\end{figure}

@@

@@

\quad There are some very interesting features here. It essentially provides a window into users' status, connections, and interaction within the Yelp ecosystem. The figure below shows a plot of the user data. We see that the user density grows around 3.25-4.25 average rating with some traces of a bell curve regarding the number of reviews. The most significant yet hardest to process attribute is the 'friends' list.

\begin{figure}[!h]
\includegraphics[width=0.45\textwidth, center]{u_reviews_vs_stars}
\end{figure}

\quad
\section{Next Steps}

\quad It is still necessary to reduce dimensionality. We?ve currently been working on the whole dataset and simply ignoring all attributes that fall in the category ?other.? This is cumbersome and it will become increasingly detrimental as we use more advanced techniques like clustering. The most prioritized attribute is the attributes portion. This attribute has the most issues because it is a miscellaneous catch-all where any attribute that didn?t fit with the others is thrown in. The issue is that it totally lacks consistency in the number and types of miscellaneous attributes included for each business. One business may have a Free Wifi attribute while others may not. After looking through the data it was clear that there wasn't one attribute in this misc. category that was reliably present for all businesses. Although we have moved on to work with the other attributes, this one still needs work to decide if it should be omitted completely or if default values should be set instead.

@@

>>>>

\quad Yelp is a website which allows users to create an account and review businesses. Our goal is to analyze data from Yelp to gain further insight for business owners to develop a more satisfactory experience for their customers. Yelp has a large public dataset available on their website with information about users, businesses, and reviews.

\quad Our first objective is to determine if certain attributes are indicative of business success with respect to their average review rating (see sec. 2.3 for revenue relation). For the businesses with many reviews, time-series data of the average star rating of the business mined from user reviews can be analyzed with ML models like Hidden Markov Models or recurrent neural networks in order to learn the likelihood of a business failing or succeeding in the future based on its current ratings. Furthermore, sentiment analysis can be used on language of reviews and matched with the star reviews to provide more information on a business's chances of success. Additionally, information about the current state of a business and likely future states could be provided to businesses to permit them to change their current practices and focus to increase their chances of success.

\quad Another way sentiment analysis will prove helpful in this project is to analyze the emotions conveyed by the reviewer as being pleased or angry at their experience with the business. Using the results of this analysis we hope to be able to determine if certain times of day, city locations, or other factors contribute to the experience of a customer. This information could also contribute to a business model. As an example, if certain locations are more prone to angry customers it could indicate that the people in that area have high standards, the businesses are poorly run, or just that those customers tend to leave angry reviews. This type of information will be less useful to businesses, but it could prove very helpful for customers or advertisers to know that businesses in a certain area tend to be reviewed much higher or lower than the average business in cities of a comparable size.

\quad Our last objective is to find recurring problems in businesses. With natural language processing tools, we can analyze review text to find the specific issues that people are having with businesses via keywords or repeated references to specific aspects of the business. Again, this will be most helpful in analyzing reviews for businesses with many reviews. If one of these businesses has many reviews which all indicate a similar issue, then that is useful information that a business owner would want to be aware of so that they can eliminate the problem, if possible.

@@

A previous winner, Hood et al., published a paper on assessing a business's current rating/review state and future expectations with respect to review numbers on Yelp. The bulk of their work centered around the creation of additional features for the data (e.g. number of similar businesses within 1km, features of subsets of reviews such as average rating and number of unique users). From there they utilized various feature reduction and clustering techniques to build a prediction model. They were able to identify features that offered greater accuracy in prediction of interest in the business (calculated as the predicted number of reviews within 6 months after a target date) as well as a create model for prediction that outperformed a basic linear regression.

@@

\quad The dataset has many optional attributes. There is an attribute simply called 'attributes' that is a shopping list of optional fields for each business. This part of the dataset is inconsistent and the same fields cannot be relied upon to be present for each business. For instance, one of these attributes is 'Bike Parking.' This attribute is either true, false, or nonexistent. In the cases when it is nonexistent, analysis based upon this attribute is obviously unreliable. Most of these miscellaneous attributes are present for much less than a majority of the data, so any mining based upon them would have to first drastically trim the data available for analysis by eliminating more than half the data. For this reason, we have chosen to eliminate the miscellaneous category rather than most of our data.

\quad In addition, the data is not only restaurants but many kinds of businesses. A datum associated with each business attempts to classify the business as a member of a category, such as 'restaurants,' 'beaches,' or 'shopping.' However some categories are extremely broad and the same type of business can be described in different ways. One example is that a bakery may be classified as belonging to 'bakeries,' but also generically as 'food.' Conversely, many businesses sell food, but the generic category 'food' will not be in their list of categories and instead the business will be labelled with more specific categories like 'Italian food,' diminishing the usefulness of these categories.

\quad The dataset provided by the Yelp organizers is fairly clean already, and the majority of the data cleaning process will be involved in handling missing data. It is possible to fill in these missing values with default values like 'false', or 'none' but it can be dangerous to make assumptions about what businesses have and don't have - for example, filling in a 'gluten-free' attribute of a business may influence our analysis of its reviews and, depending on exposure, maybe even mislead Yelp users as to what the business provides.

\quad The Yelp dataset contains up to a million different attributes about businesses from 11 different cities in 4 countries. The majority of these attributes will be unnecessary so there will be a large effort in data reduction. For instance, the dataset includes photographs of the exterior and interior of some businesses (also many images of various kinds of food served by the businesses), but our questions does not require image analysis. There also redundant attributes such as postal codes that are unnecessary because latitude and longitude is more accurate.

@@

\quad The Yelp dataset comes from the Yelp Dataset Challenge. This is an ongoing challenge and this is the ninth round. The challenge offers \$5,000 to the winner and the  final submission is an academic paper.

@@

\quad The tools and methods used for this project will no longer include a database for storing or integrating the data. Due to difficulties described in section 6 and relating to the unknown JSON variant used by the preparers of the Yelp dataset, storing the data in a database is not feasible at this time. Given that the data is already quite clean save for the format, with minimal replacement of missing values and some other changes necessary, it is possible for each group member to have a copy of the data stored locally that may be processed with our tools, with smaller subsets of data, results, and scripts shared via GitHub.

\quad The datasets will be merged locally, and while this does increase the amount of computational work necessary for each member?' machine, the dataset is still of a small enough size (less than 10 GB, excluding the photos) that this is practical. We will still use various Python libraries and modules for cleaning and preprocessing, and various other tools for transformation, mining, pattern evaluation, and knowledge presentation. The lack of a centralized database in removes the need for a number of Python libraries like Dora and FTFY ('fixes text for you') intended for data cleaning and preprocessing, but increases the necessity of group communication to reduce duplication of work and miscommunications. We currently use and will continue to use Slack, email, and in-person meetings to communicate and synchronize our efforts.

\quad Numpy, Scipy, and Pandas will be used to transform and mine the data, and other Python modules like Seaborn and Plotly (a Python API for D3) will be used for knowledge visualization and presentation. Rather than relying on tools like Weka, the Text Analytics API from Microsoft Cognitive Services, or the Facebook fastText pre-trained word vectors, we will instead use Python's Natural Language Toolkit for sentiment analysis and other language processing.

@@

\quad Also, evaluating the sentiment analysis of the review text will likely be simple given that we will be using tools like NLTK (see following section) to perform sentiment analysis rather than creating entirely custom algorithms. These tools incorporate helpful metrics to analyze the error of a particular output. Alternatively, we can manually verify the sentiment of a (relatively small) number of reviews by reading them and comparing our human judgements of the sentiment to the output of the tools. For evaluating recurring problems, we will employ a similar approach, except we will group reviews by business and analyze them as a time series to and recurring issues. For recurring problems specifically, the most expedient method of evaluation may well be to read a small number of reviews and manually determine evaluate the performance of our tools in detecting recurrent issues.

\quad For all three of the objective questions, employing association rule frameworks that use support-confidence frameworks will be helpful to find and measure strong associations, while null-invariant measures like the Kulczynski measure or the cosine measure will be helpful to determine how interesting these associations are.

@@

\quad There have been several major changes to the project tools and methods since the project proposal. Initially we planned on using a relational database to store, integrate, and access the data, but but we encountered many difficulties in just uploading the data. The inconsistencies between the user and business data made it exceedingly difficult to successfully import the data en masse into the database, so instead we decided to import the data straight into Weka. However, the format of the JSON file also did not allow a successful import into Weka.

\quad Next we converted the JSON file into CSV format using a Python script and tried to upload it to Weka again. Once again Weka had issues due to the mismatch in attributes for each business (some had more attributes while others had less, likely due to the ?miscellaneous? category mentioned in section 3.1).

\quad Instead we opted to use a different tool, RapidMiner. Like Weka, RapidMiner incorporates extensive machine learning functionality like classification, regression, clustering, and association and frequent itemset mining. RapidMiner is commercial software, but an educational version is available that removes certain limitations on the free version like only being able to process 10,000 rows at a time. Additionally, RapidMiner is much more compatible with the missing data in our dataset and automatically corrects the formatting issues we had when attempting to use Weka. It does not support JSON so we reused the conversion code (found on the team GitHub as convert.py) to convert the data files to CSV.

@@

\quad After importing the CSV formatted version of the data we tested some basic visualization of the data. For instance, in Figure 1 we created a histogram of the average star for a business versus the state in which it's located. Basic graphs like these help refine our question. For example, why does Illinois have significantly lower average stars than others?

@@

\begin{figure}[h]
\includegraphics[width=0.5\textwidth, center]{b_rating_by_state}
\label{fig:Average star rating by state}
\caption{Note the outliers IL and NTH}
\end{figure}

@@

\begin{figure}[!h]
\includegraphics[width=0.5\textwidth, center]{3d_businesses_lat_long}
\caption{Clusters likely indicate areas of high population}
\label{fig:Review count by latitude and longitude}
\end{figure}

@@

\subsection{Reviews}

\quad One of the major goals of the project is to find recurrent problems and other issues with the businesses, and this was to be done by grouping reviews by business and then analyzing trends in the language used in the reviews over time. However, given that we were unable to find a suitable workaround for our difficulties with Weka, we used sentiment analysis tools from the Natural Language Toolkit (NLTK) to analyze the review language instead. NLTK is a language-processing and analysis platform that uses Python, and a good replacement for the language processing capabilities of Weka.

\quad Using a custom script (see GitHub; nltk\_review\_subset\_process.py) to analyze the sentiment of the first 100,000 reviews (about 2.5\% of the available data), we discovered that the average sentiment score among all analyzed reviews was around 0.113, and there were 17295 negative reviews compared to 76968 positive reviews. The remaining 5737 reviews had composite sentiment scores within -0.2 < score < 0.2, so they were excluded from further analysis since they were neutral. This means that the average Yelp review (at least within the analyzed subset) is positive rather than negative or neutral in sentiment, perhaps indicating that Yelp users are more inclined to write reviews of businesses with which they had a good experience. Only around 2.5\% of the data is currently processed by the file, because processing the entire file is prohibitively expensive.

@@

\quad There are some very interesting features here. It essentially provides a window into users' status, connections, and interaction within the Yelp ecosystem. The figure below shows a plot of the user data. We see that the user density grows around 3.25-4.25 average rating with some traces of a bell curve regarding the number of reviews. In face, it appears that the user data skews so heavily in this grouping that all other measures

\begin{figure}[!h]
\includegraphics[width=0.5\textwidth, center]{u_reviews_vs_stars}
  \caption{Note how the density increases between 3.25 and 4.25 starts}
  \label{fig:Plot of review count vs average star rating}
\end{figure}

\quad One of the most significant yet hardest to process attributes is the 'friends' list. This attribute provides an opportunity to model our data as a graph. Such a network could provide very relevant clustering through sparseness measures and cutting in addition to accounting for user/business influence or finding deeper trends according to network connectivity.  
\quad
\section{Next Steps}

\quad It is still necessary to reduce dimensionality. We've currently been working on the whole dataset and simply ignoring all attributes that fall in the category 'other.' This is cumbersome and it will become increasingly detrimental as we use more advanced techniques like clustering. The most prioritized attribute is the attributes portion. This attribute has the most issues because it is a miscellaneous catch-all where any attribute that didn?t fit with the others is thrown in. The issue is that it totally lacks consistency in the number and types of miscellaneous attributes included for each business. One business may have a Free Wifi attribute while others may not. After looking through the data it was clear that there wasn?t one attribute in this misc. category that was reliably present for all businesses. Although we have moved on to work with the other attributes, this one still needs work to decide if it should be omitted completely or if default values should be set instead.

@@

\quad The user and review datasets are the largest we have. We would like to use these to find useful information about the people and the reviews they give. Because Yelp is a platform on which anybody can create an account and write a review, it is possible that not everybody is honest and fair in their reviews. Furthermore, different people may assign scores for different reasons.

\quad Some useful attributes in the user data alone are the number of reviews that an individual has made, the length of time a user has been active on Yelp, and whether their reviews tend to be tagged as useful by other users.

\quad One additional consideration is processing power. Our chosen tool seems to be very computationally intensive, so we will very likely need to use sampled data to test hypotheses then turn to python scripts to actually run over the entirety of our training, testing, and validations sets. Furthermore, our data is currently only linked through review, user, and business id pointers. While there?s a great deal to be learned from each collection individually, we will need to connect our disparate sets by these terms.

\quad Currently, the review dataset is too large to process in detail. In the future, we will continue taking subset samples of a more reasonable size and processing those instead, as well as increasing the complexity of our analysis. Combining the review dataset with the business dataset (using the ?business\_id? field as a key) will enable us to produce maps of where the largest number of reviews with negative and positive sentiment are, as well as seeing where users leave the most strongly-worded reviews. We hope that additional advanced sentiment analysis will be useful in determining underlying factors that influence a business's success, as well as in identifying recurrent problems.

\quad Once we have our sentiment analysis we can combine this data with our user data and business data to find patterns in the sentiment of reviews. We would like to compare the sentiment of reviews to the area they are coming from, the businesses they apply to, and whether reviews of a given sentiment are considered beneficial by other users. Using this information we can then narrow down the factors that are most influential to the score of a business in order to draw our final conclusions.
